IOIs Analysis: The Sound of Teaching Music 2

### 2020-01-06 script/RNotebook/ioi_1.Rmd
add stats
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of all analyses for a tempo.

# Outline - Interonset intervals (IOIs)
### [Experimental Design](#design)

### [Articulation](#ioi-articulation)

- Individual results
  + [All](#art_ind_all)
  + [Phrase bondaries](#art_ind_phrase)
  + [Trial by trial](#art_ind_trial)
  + [Variability all](#art_ind_var_all)
  + [Variability trial-by-trial](#art_ind_var_trial)

- Group results and stats (if available)
  + [All](#art_group_all)
  + [Phrase bondaries](#art_group_phrase)
  + [Trial by trial](#art_group_trial)
  + [Variability all](#art_group_var_all)
  + [Variability trial-by-trial](#art_group_var_trial)

- Order analysis (teaching/performing first?)
  + [All](#art_order_group_all)

- [Sequence plot](#sequence_art)

### [Dynamics](#ioi-dynamics)

- Individual results
  + [All](#dyn_ind_all)
  + [Phrase bondaries](#dyn_ind_phrase)
  + [Trial by trial](#dyn_ind_trial)
  + [Variability all](#dyn_ind_var_all)
  + [Variability trial-by-trial](#dyn_ind_var_trial)

- Group results and stats (if available)
  + [All](#dyn_group_all)
  + [Phrase bondaries](#dyn_group_phrase)
  + [Trial by trial](#dyn_group_trial)
  + [Variability all](#dyn_group_var_all)
  + [Variability trial-by-trial](#dyn_group_var_trial)

- Order analysis (teaching/performing first?)
  + [All](#dyn_order_group_all)

- [Sequence plot](#sequence_dyn)

@@

```{r extract, include = FALSE}
df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv

@@

#### Plot
##### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_art_all_bar, echo = FALSE}
p_ioi_art_all <- ggplot(data = ioi_art_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
  theme(legend.position = "None")
p_ioi_art_all
```

```{r ioi_art_all_box, echo = FALSE}
p_ioi_art_all_box <- ggplot(data = ioi_art, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + #coord_cartesian(ylim = c(100, 210)) +
  theme(legend.position = "None")
p_ioi_art_all_box
```

@@

### 3. Phrase boundaries - individual{#art_ind_phrase}
- Compare IOIs at phrase boundaries with others fir each participant.

```{r ioi_art_ch, echo = FALSE}
# Define phrase boundaries
df_ioi_art$Boundary <- "No"
df_ioi_art$Boundary[df_ioi_art$Interval == 8 | df_ioi_art$Interval == 16 | df_ioi_art$Interval == 24 | df_ioi_art$Interval == 41 | df_ioi_art$Interval == 49 | df_ioi_art$Interval == 57] <- "Yes"

# For each individual
ioi_art_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_ch <- cbind(ioi_art_ch[,1:4], as.data.frame(ioi_art_ch[,5]))
# Change colnames
colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
print(ioi_art_ch)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_art_ch_bar, echo = FALSE}
p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
p_ioi_art_ch
```

### 4. Phrase boundaries - group{#art_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.

@@

#### Plot
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_art_ch_all_bar, echo = FALSE}
p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  #facet_grid(. ~ Boundary) +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 250))
p_ioi_art_ch_all_bar
```

```{r ioi_art_ch_all_box, echo = FALSE}
p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation")
p_ioi_art_ch_box
```

@@

##### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
ioi_art_ch_aov <- ezANOVA(
  data = df_ioi_art
  , dv = .(IOI)
  , wid = .(SubNr)
  , within = .(Condition, Boundary)
  , type = 3
  , detailed = TRUE
)
ioi_art_ch_aov$ANOVA$F <- round(ioi_art_ch_aov$ANOVA$F, 4)
ioi_art_ch_aov$ANOVA$p <- round(ioi_art_ch_aov$ANOVA$p, 4)
print(ioi_art_ch_aov$ANOVA)
```

@@

### 5. Trial by trial - individual{#art_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi_art_trial, echo = FALSE}
# For each individual
ioi_art_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_art,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_trial <- cbind(ioi_art_trial[,1:4], as.data.frame(ioi_art_trial[,5]))
# Change colnames
colnames(ioi_art_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
print(ioi_art_trial)
```

#### Plot
##### Line graph
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_art_trial_line, echo = FALSE, fig.height = 8}
p_ioi_art_trial <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial
```

### 6. Trial by trial - group{#art_group_trial}
- Average IOIs for each trial for each condition.

@@

#### Plot
##### Line graph
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_art_trial_all_line, echo = FALSE}
p_ioi_art_trial_all <- ggplot(data = ioi_art_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial_all
```

### 7. Variability all - individual{#art_ind_var_all}
- Average variability for each participant.
- IOI variability = SD/mean IOI (per trial)

@@

#### Statistics
##### Normality check
```{r normality3, echo = FALSE}
ioi_art_var_norm <- by(ioi_art_var$Mean, list(ioi_art_var$Condition), shapiro.test)
print(ioi_art_var_norm)
```

@@

```{r ioi_dyn_all_box, echo = FALSE}
p_ioi_dyn_all_box <- ggplot(data = ioi_dyn, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + #coord_cartesian(ylim = c(100, 210)) +
  theme(legend.position = "None")
p_ioi_dyn_all_box
```

@@

### 3. Phrase boundaries - individual{#dyn_ind_phrase}
- Compare IOIs at phrase boundaries with others fir each participant.

```{r ioi_dyn_ch, echo = FALSE}
# Define phrase boundaries
df_ioi_dyn$Boundary <- "No"
df_ioi_dyn$Boundary[df_ioi_dyn$Interval == 8 | df_ioi_dyn$Interval == 16 | df_ioi_dyn$Interval == 24 | df_ioi_dyn$Interval == 41 | df_ioi_dyn$Interval == 49 | df_ioi_dyn$Interval == 57] <- "Yes"

# For each individual
ioi_dyn_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_dyn,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_dyn_ch <- cbind(ioi_dyn_ch[,1:4], as.data.frame(ioi_dyn_ch[,5]))
# Change colnames
colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
print(ioi_dyn_ch)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_dyn_ch_bar, echo = FALSE}
p_ioi_dyn_ch <- ggplot(data = ioi_dyn_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
p_ioi_dyn_ch
```

### 4. Phrase boundaries - group{#dyn_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.

@@

#### Plot
##### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_dyn_ch_all_bar, echo = FALSE}
p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  #facet_grid(. ~ Boundary) +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 250))
p_ioi_dyn_ch_all_bar
```

```{r ioi_dyn_ch_all_box, echo = FALSE}
p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics")
p_ioi_dyn_ch_box
```

#### Statistics
##### Normality check
```{r normality5, echo = FALSE}
ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
print(ioi_dyn_ch_norm)
```

@@

### 5. Trial by trial - individual{#dyn_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi_dyn_trial, echo = FALSE}
# For each individual
ioi_dyn_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_dyn_trial <- cbind(ioi_dyn_trial[,1:4], as.data.frame(ioi_dyn_trial[,5]))
# Change colnames
colnames(ioi_dyn_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
print(ioi_dyn_trial)
```

#### Plot
##### Line graph
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_dyn_trial_line, echo = FALSE, fig.height = 8}
p_ioi_dyn_trial <- ggplot(data = ioi_dyn_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial
```

### 6. Trial by trial - group{#dyn_group_trial}
- Average IOIs for each trial for each condition.

@@

#### Plot
##### Line graph
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_dyn_trial_all_line, echo = FALSE}
p_ioi_dyn_trial_all <- ggplot(data = ioi_dyn_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial_all
```

@@

#### Statistics
##### Normality check
```{r normality6, echo = FALSE}
ioi_dyn_var_norm <- by(ioi_dyn_var$Mean, list(ioi_dyn_var$Condition), shapiro.test)
print(ioi_dyn_var_norm)
```

@@

## IOI-Order-Articulation
- Investigate whether there is any order effect of the conditions.

### 1. All{#art_order_group_all}
```{r ioi_art_order_all, echo = FALSE}
# For each participant
ioi_art_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_art,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
ioi_art_order <- cbind(ioi_art_order[,1:4], as.data.frame(ioi_art_order[,5]))
# Change colnames
colnames(ioi_art_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD")

# Group mean
ioi_art_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_art_order,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_order_all <- cbind(ioi_art_order_all[,1:3], as.data.frame(ioi_art_order_all[,4]))
# Change colnames
colnames(ioi_art_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM")
print(ioi_art_order_all)
```

### Plot
#### Bar graph
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)

```{r ioi_art_order_all_bar, echo = FALSE}
p_ioi_art_order_all <- ggplot(data = ioi_art_order_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge()) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(Order ~ .) +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
  theme(legend.position = "None")
p_ioi_art_order_all
```

>>>>

---
title: 'IOIs Analysis: The Sound of Teaching Music 2'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of all analyses for a tempo (IOIs).
  - data <- data_ioi_1.csv

@@

```{r extract, include = FALSE}
df_ioi <- read.csv("./trimmed/data_ioi_1.csv", header = T, sep = ",", dec = ".") # read a trimmed csv

@@

#### Plot
##### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the ideal tempo

```{r ioi_art_all_bar, echo = FALSE}
p_ioi_art_all <- ggplot(data = ioi_art_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(.75, 1.1)) +
  theme(legend.position = "None")
p_ioi_art_all
```

##### Box plot
- A dashed line represents the ideal tempo

```{r ioi_art_all_box, echo = FALSE}
p_ioi_art_all_box <- ggplot(data = ioi_art, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Articulation") +
  theme(legend.position = "None")
p_ioi_art_all_box
```

@@

### 3. Subcomponent change - individual{#art_ind_phrase}
- Compare normIOIs at which Subcomponent changes (e.g., Legato to Staccato) with others fir each participant.

```{r ioi_art_ch, echo = FALSE}
# Define phrase boundaries
df_ioi_art$Boundary <- "No"
df_ioi_art$Boundary[df_ioi_dyn$Subcomponent == "LtoS" | df_ioi_dyn$Subcomponent == "StoL" | df_ioi_dyn$Subcomponent == "FtoP" | df_ioi_art$Subcomponent == "PtoF"] <- "Yes"

# For each individual
ioi_art_ch <- aggregate(normIOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_ch <- cbind(ioi_art_ch[,1:4], as.data.frame(ioi_art_ch[,5]))
# Change colnames
colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
print(ioi_art_ch)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the ideal tempo

```{r ioi_art_ch_bar, echo = FALSE}
p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Articulation")
p_ioi_art_ch
```

### 4. Subcomponent change - group{#art_group_phrase}
- Compare IOIs at which Subcomponent changes (e.g., Legato to Staccato) with others in the two conditions.
- Hypothesis: Participants will play more slowly where Subcomponent changes compared with other locations.

@@

#### Plot
- Error bars show SEM of the mean
- A dashed line represents the ideal tempo

```{r ioi_art_ch_all_bar, echo = FALSE}
p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(.75, 1.1))
p_ioi_art_ch_all_bar
```

##### Box plot
- A dashed line represents the ideal tempo

```{r ioi_art_ch_all_box, echo = FALSE}
p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Articulation")
p_ioi_art_ch_box
```

@@

##### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
ioi_art_ch_aov <- ezANOVA(
  data = df_ioi_art
  , dv = .(normIOI)
  , wid = .(SubNr)
  , within = .(Condition, Boundary)
  , type = 3
  , detailed = TRUE
)
ioi_art_ch_aov$ANOVA$F <- round(ioi_art_ch_aov$ANOVA$F, 4)
ioi_art_ch_aov$ANOVA$p <- round(ioi_art_ch_aov$ANOVA$p, 4)
print(ioi_art_ch_aov$ANOVA)
```

@@

### 5. Trial by trial - individual{#art_ind_trial}
- Average normIOIs for each trial for each participant.

```{r ioi_art_trial, echo = FALSE}
# For each individual
ioi_art_trial <- aggregate(normIOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_art,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_trial <- cbind(ioi_art_trial[,1:4], as.data.frame(ioi_art_trial[,5]))
# Change colnames
colnames(ioi_art_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
print(ioi_art_trial)
```

#### Plot
##### Line graph
- A dashed line represents the ideal tempo

```{r ioi_art_trial_line, echo = FALSE, fig.height = 8}
p_ioi_art_trial <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Normalised IOI", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial
```

### 6. Trial by trial - group{#art_group_trial}
- Average normIOIs for each trial for each condition.

@@

```{r ioi_art_trial_all_line, echo = FALSE}
p_ioi_art_trial_all <- ggplot(data = ioi_art_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Normalised IOI", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial_all
```

### 7. Variability all - individual{#art_ind_var_all}
- Average variability for each participant.
- normIOI variability = SD/mean IOI (per trial)

@@

#### Statistics
##### Normality check
```{r normality3, echo = FALSE}
ioi_art_var_norm <- by(ioi_art_var$Mean, list(ioi_art_var$Condition), shapiro.test)
print(ioi_art_var_norm)
```
- Violate the normality assumption

@@

```{r ioi_dyn_all_box, echo = FALSE}
p_ioi_dyn_all_box <- ggplot(data = ioi_dyn, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics") + 
  theme(legend.position = "None")
p_ioi_dyn_all_box
```

@@

### 3. Subcomponent changes - individual{#dyn_ind_phrase}
- Compare normIOIs at which Subcomponent changes (e.g., Forte to Piano) with others fir each participant.

```{r ioi_dyn_ch, echo = FALSE}
df_ioi_dyn$Boundary <- "No"
df_ioi_dyn$Boundary[df_ioi_dyn$Subcomponent == "LtoS" | df_ioi_dyn$Subcomponent == "StoL" | df_ioi_dyn$Subcomponent == "FtoP" | df_ioi_dyn$Subcomponent == "PtoF"] <- "Yes"

# For each individual
ioi_dyn_ch <- aggregate(normIOI~SubNr*Condition*Skill*Boundary, data = df_ioi_dyn,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_dyn_ch <- cbind(ioi_dyn_ch[,1:4], as.data.frame(ioi_dyn_ch[,5]))
# Change colnames
colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
print(ioi_dyn_ch)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the ideal tempo

```{r ioi_dyn_ch_bar, echo = FALSE}
p_ioi_dyn_ch <- ggplot(data = ioi_dyn_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics")
p_ioi_dyn_ch
```

### 4. Subcomponent changes - group{#dyn_group_phrase}
- Compare IOIs at which Subcomponent changes with others in the two conditions.
- Hypothesis: Participants will play more slowly where Subcomponent changes compared with other locations.

@@

#### Plot
##### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the ideal tempo

```{r ioi_dyn_ch_all_bar, echo = FALSE}
p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(.75, 1.1))
p_ioi_dyn_ch_all_bar
```

##### Box plot
- A dashed line represents the ideal tempo

```{r ioi_dyn_ch_all_box, echo = FALSE}
p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "normIOI: Dynamics")
p_ioi_dyn_ch_box
```

#### Statistics
##### Normality check
```{r normality5, echo = FALSE}
ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
print(ioi_dyn_ch_norm)
```
  - Violate the normality assumption

@@

### 5. Trial by trial - individual{#dyn_ind_trial}
- Average normIOIs for each trial for each participant.

```{r ioi_dyn_trial, echo = FALSE}
# For each individual
ioi_dyn_trial <- aggregate(normIOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_dyn_trial <- cbind(ioi_dyn_trial[,1:4], as.data.frame(ioi_dyn_trial[,5]))
# Change colnames
colnames(ioi_dyn_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
print(ioi_dyn_trial)
```

#### Plot
##### Line graph
- A dashed line represents the ideal tempo.

```{r ioi_dyn_trial_line, echo = FALSE, fig.height = 8}
p_ioi_dyn_trial <- ggplot(data = ioi_dyn_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Normalised IOI", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial
```

### 6. Trial by trial - group{#dyn_group_trial}
- Average normIOIs for each trial for each condition.

@@

```{r ioi_dyn_trial_all_line, echo = FALSE}
p_ioi_dyn_trial_all <- ggplot(data = ioi_dyn_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Normalised IOI", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial_all
```

@@

#### Statistics
##### Normality check
```{r normality6, echo = FALSE}
ioi_dyn_var_norm <- by(ioi_dyn_var$Mean, list(ioi_dyn_var$Condition), shapiro.test)
print(ioi_dyn_var_norm)
```
  - Violate the normality assumption

@@

## Sequence plot
#### Articulation{#sequence_art}
```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
# For each individual
ioi_art_seq <- aggregate(normIOI~SubNr*Condition*Skill*Interval, data = df_ioi_art,
                                 FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_seq <- cbind(ioi_art_seq[,1:4], as.data.frame(ioi_art_seq[,5]))
# Change colnames
colnames(ioi_art_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
